---
title: "MutateRate"
author: Skye Antinozzi and Mitchell Finzel
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knapPI_20_allP <- read.csv("../Labs/AI/simple-search/data/knapPI_20_allP", sep="")
knapPI_20_allP$non_neg = ifelse(knapPI_20_allP$X1320<0,0,knapPI_20_allP$X1320)
```
<!--knapPI_200_allP <- read.csv("../Labs/AI/simple-search/data/knapPI_200_allP", sep="")
knapPI_200_allP$non_neg = ifelse(knapPI_200_allP$X0.1<0,0,knapPI_200_allP$X0.1)-->


##Setup
We chose to look at how mutation rate affects score by providing mutation rate as a commandline argument. This example uses knapsack item sets of 20 with percentages of the total item size 5,10,25,50 and 100 percents.

---

```{r pressure, echo=FALSE}
boxplot(knapPI_20_allP$non_neg ~ knapPI_20_allP$X1, names=c("5%", "10%", "25%", "50%", "100%"),main="20 Items",xlab="Mutation Rate (% of items)", ylab="Score")
```

---

##Results
As can be seen in the boxplots the mutation rate didn't have much effect between 5% and 25%. At 50% however a large drop in mean score occurs with 100% mutation rate, essentially randomized, having no score above 0. At 50% of the item set size the score begins to approach 0, because this is penalized score the scores are becoming more negative.


##Future Work
We would like to test a finer mutation rate gradient between 25% and 100% to see when the decline in score begins to slope more harshly. We would like to replicate our process with larger item set knapsack problems as well.

<!--```{r,echo=FALSE}
boxplot(knapPI_200_allP$X0.1 ~ knapPI_200_allP$X10, names=c("5%", "10%", "25%", "50%", "100%"), main="200 Items")
```-->

